Multi-stream portrait of the Cosmic web

### 2018-06-06 Draft/Chapter3/Paper2015.tex
may 21
thesis update

\chapter{Multi-stream portrait of the Cosmic web}

\section{Introduction} \label{sec:intro}

@@

Large differences in the estimates of volume and mass fractions made by different groups are not surprising if we recognize
considerable differences in the definitions of the components of the cosmic web and numerical methods used in the estimates. Without trying to provide an exhaustive review of all definitions and techniques used for the quantitative morphological 
analysis of the web  we just briefly describe a few approaches in order to illustrate how different they could be. 
Some groups study the web morphology using only coordinates of simulation particles, while others use the particle velocities too. 
Transforming data from  point sets to the density and other fields on a grid is also often used
because fields allow to use a variety of mathematical techniques not available for the particle sets.
However this step can be done by a variety of methods some as simple as cloud-in-cell (CIC), or more complicated as smoothed-particle hydrodynamics (SPH), 
or using Voronoi and Delaunay tessellations
as in the Delaunay Tessellation Field Estimator (DTFE) method (~\citealt{Weygaert2009}, \citealt{Cautun2014a}).  
Recently a new method called a discrete persistent structure extractor (DisPerSE \citealt{Sousbie2011a},  \citealt{Sousbie2011b} allowing to identify haloes and other components of the web directly from the particles 
has been designed.  This method can be applied to the galaxy catalogues, for instance \citet{Sousbie2011b} applied it to SDSS catalogue and extracted the filaments (which are made available online).

@@

Using the full six-dimensional information allows one to generate new fields which  provide 
additional useful information about the evolution and morphology of the  web. 
One of them is a multi-stream field in Eulerian space, which will be the focus of this paper. 
Another example is the flip-flop field in Lagrangian space. In cosmological context it was first used in the ZA.
\citet{Vogelsberger2011} used it in a study of multi-stream structure of galaxy size haloes.
\citet{Shandarin2014} applied it for identifying subhaloes  in dark matter haloes.
A similar although somewhat simplistic realization of this idea has been revealed in the ORIGAMI method
used for the analysis of the web ( \citealt{Falck2012},  \citealt{Falck2015} ).
 Although these fields cannot be used directly on observational data because the full phase-space information
 is not available, they provide much deeper insight into non-linear clustering of collision-less dark matter
and reveal new features of the web.

@@

We have utilized the data from cosmological N-body simulations by Gadget-2 \citep{Springel2005} for 100 $h^{-1}$ Mpc  and 200 $h^{-1}$ Mpc box 
sizes with $128^3$, $256^3$ and $512^3$ grids. Each particle is between  $10^9 - 10^{12} M_{\odot}$. 
The initial conditions and cosmological parameters are consistent with the Planck cosmology. We utilize the initial Lagrangian box and do a three-dimensional mapping onto corresponding evolved simulations. In addition, for local multi-stream analyses around haloes, we have utilized halo catalogues for each of these simulation boxes. These haloes are detected using FOF method considering objects with more than 20 particles found at linking length, $ b= 0.2$.  
%-------------------------------------------------------------------------------------------------------------------------------
\section{Multi-stream field calculation}
\label{subsec:method}

@@

\begin{table}
  \caption{Comparison of the approximate power law dependences of curve fits in Fig. \ref{fig:Vfr_all} and Fig. \ref{fig:Mfr_all}. Power law relations for volume fraction $f_{vol}(n_{str})$ and mass fractions $f_{mass} (n_{str}) $ as a function of number of streams at $n_{str} \geq 5$ are shown (amplitudes are not shown). The boxes of size 100 $h^{-1}$ Mpc on  $128^{3}$ grids and, 200 $h^{-1}$ Mpc on  $256^{3}$ grids have same  $L/N = 0.78 h^{-1}$ Mpc. Similarly, $L/N = 0.39 h^{-1}$ Mpc for boxes of size 100 $h^{-1}$ Mpc on  $256^{3}$ grids and, 200 $h^{-1}$ Mpc on  $512^{3}$ grids.}
\begin{tabular}{|l|r|r|r|r|}
\hline
%\multicolumn{1}{|c|}{} & \multicolumn{4}{l|}{Multi-stream analysis (This work)} & \multicolumn{4}{c|}{ORIGAMI \citep{Falck_Neyrinck:14}}    \\ \hline
$L/N$                & $0.19$ & $0.39$& $0.78$ & $1.56$  \\ \hline
$f_{vol} (n_{str}) $ Vs. $n_{str}$                 & $-2.1$   & $-2.3$    & $-2.5$      & $-2.9$  \\ \hline
$f_{mass} (n_{str}) $   Vs. $n_{str}$            & $-1.1$   & $-1.2$   & $-1.4$       & $-2.0$     \\ \hline
\end{tabular}
 \label{tab:Compare_Slopes}
\end{table}

 \begin{table}
  \caption{Comparison of the volume and mass fractions of the void ($n_{str} = 1$) regions of the cosmic web for various simulation boxes at refinement factor of 1. Mean density is the ratio of mass fraction to the volume fraction. It is given in units of the mean density of the universe. The boxes of size 100 $h^{-1}$ Mpc on  $128^{3}$ grids and, 200 $h^{-1}$ Mpc on  $256^{3}$ grids have same  $L/N = 0.78 h^{-1}$ Mpc. Similarly, $L/N = 0.39 h^{-1}$ Mpc for boxes of size 100 $h^{-1}$ Mpc on  $256^{3}$ grids and, 200 $h^{-1}$ Mpc on  $512^{3}$ grids.}
\begin{tabular}{|l|r|r|r|r|}
\hline
%\multicolumn{1}{|c|}{} & \multicolumn{4}{l|}{Multi-stream analysis (This work)} & \multicolumn{4}{c|}{ORIGAMI \citep{Falck_Neyrinck:14}}    \\ \hline
$L/N$                & $0.19$ & $0.39$& $0.78$ & $1.56$  \\ \hline
Volume Fraction (\%)                 & $88$   & $90$    & $93$      & $96$  \\ \hline
Mass Fraction   (\%)               & $24$   & $36$   & $55$       & $77$     \\ \hline
Mean density                    & $0.27$ & $0.40$    & $0.59$        & $0.80$   \\ \hline
\end{tabular}
 \label{tab:Compare_LN}
\end{table}

 \begin{table}
  \caption{Comparison of the volume and mass fractions of the void ($n_{str} = 1$) regions of the cosmic web for a simulation box at different refinement factors. Mean density is the ratio of mass fraction to the volume fraction. All the multi-streams for simulation box of length 100 $h^{-1}$ Mpc on raw resolution of $128^{3}$ grids ($ L/N = 0.78 h^{-1}$) }
\begin{tabular}{|l|r|r|r|}
\hline
%\multicolumn{1}{|c|}{} & \multicolumn{4}{l|}{Multi-stream analysis (This work)} & \multicolumn{4}{c|}{ORIGAMI \citep{Falck_Neyrinck:14}}    \\ \hline
Refinement factor               & $1$ & $2$& $8$   \\ \hline
Volume Fraction (\%)                 & $93$   & $93$    & $93$       \\ \hline
Mass Fraction   (\%)               & $55$   & $44$   & $32$           \\ \hline
Mean density                    & $0.59$ & $0.47$    & $0.35$        \\ \hline
\end{tabular}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=10.cm]{Chapter3/Source_v2/fig10a}
\includegraphics[width=10.cm]{Chapter3/Source_v2/fig10b}
\includegraphics[width=7.cm]{Chapter3/Source_v2/fig10c}
\end{minipage}\hfill
\caption{Halo of mass $1.1 \times 10^{12} M_{\odot} $  and FOF radius 0.2 $h^{-1}$ Mpc (dotted-violet line). This halo just has 26 particles, hence the resolution is lesser than previous haloes for surfaces with low radii. Top: There is a bump in fraction of each of 3+ streams a little over 1 $h^{-1}$ Mpc. This is due to the presence of an additional halo nearby, as seen in the projections. The Mollweide projections from 0.4 $h^{-1}$ Mpc to 2 $h^{-1}$ Mpc have high stream flow regions near the lower surfaces of diagnostic spheres. Bottom: Corresponding FOF halo (red, at the centre) has a more massive neighbouring FOF halo (blue) within distance of 2 $h^{-1}$ Mpc. The 17+ stream regions (green) are increased around the neighbouring halo.} 
\label{fig:4322}
\end{figure}

@@

 \begin{table*}
  \caption{Comparison of the volume and mass fraction of the elements of the cosmic web between our analysis and \citet{Falck2015}. $ L/N = 0.78 h^{-1}$ Mpc for simulations used in both techniques. We use a refinement factor of 8 for the multi-stream grid. The mean density is given in units of the mean density of the universe. }
\begin{tabular}{|l|r|r|r|r|r|r|r|r|}
\hline
\multicolumn{1}{|c|}{} & \multicolumn{4}{l|}{Multi-stream analysis (This work)} & \multicolumn{4}{c|}{ORIGAMI \citep{Falck2015}}    \\ \hline
                       & Voids & Walls & Filaments & Haloes  & Voids & Walls & Filaments & Haloes \\ \hline
Volume Fraction (\%)                 & $93$   & $7$    & $< 1$      & $< 0.1$ & $84$   & $12$   & $3$        & $< 1$  \\ \hline
Mass Fraction   (\%)               & $32$   & $35$   & $17$       & $14$    & $26$   & $19$   & $19$       & $35$   \\ \hline
Mean density                    & $0.34$ & $5$    & $>17$      & $>140$  & $0.31$ & $2.2$  & $6$        & $>35$  \\ \hline
\end{tabular}
 \label{tab:Compare}
\end{table*}

@@

\section*{Acknowledgements}
The authors are grateful to Gustavo Yepes and Noam Libeskind for providing the simulations used in this study and the Sir John Templeton Foundation for the support.The authors would also like to thank the anonymous referee for careful review and useful suggestions.

>>>>

\chapter{Multi-stream portrait of the Cosmic web}\label{chapter3}

In this chapter, we report the results of the first study of the multi-stream environment of dark matter haloes in cosmological N-body simulations in the $\Lambda$CDM cosmology. We arrive at heuristic parameters for delineating walls, filaments and haloes in the multi-stream web, while single streaming regions are simply devoid of gravitational collapse, i.e., are voids. The results in the chapter were first discussed in \cite{Ramachandra2015}.

\section{Brief introduction} \label{sec:intro}

@@

Large differences in the estimates of volume and mass fractions made by different groups are not surprising if we recognize
considerable differences in the definitions of the components of the cosmic web and numerical methods used in the estimates. Without trying to provide an exhaustive review of all definitions and techniques used for the quantitative morphological 
analysis of the web  we just briefly describe a few approaches in order to illustrate how different they could be. 
Some groups study the web morphology using only coordinates of simulation particles, while others use the particle velocities too. 
Transforming data from  point sets to the density and other fields on a grid is also often used
because fields allow to use a variety of mathematical techniques not available for the particle sets.
However this step can be done by a variety of methods some as simple as cloud-in-cell (CIC), or more complicated as smoothed-particle hydrodynamics (SPH), 
or using Voronoi and Delaunay tessellations
as in the Delaunay Tessellation Field Estimator (DTFE) method (~\citealt{Weygaert2009}, \citealt{Cautun2014a}).  
Recently a new method called a discrete persistent structure extractor (DisPerSE \citealt{Sousbie2011f},  \citealt{Sousbie2011e} allowing to identify haloes and other components of the web directly from the particles 
has been designed.  This method can be applied to the galaxy catalogues, for instance \citet{Sousbie2011e} applied it to SDSS catalogue and extracted the filaments (which are made available online).

@@

Using the full six-dimensional information allows one to generate new fields which  provide 
additional useful information about the evolution and morphology of the  web. 
One of them is a multi-stream field in Eulerian space, which will be the focus of this paper. 
Another example is the flip-flop field in Lagrangian space. In cosmological context it was first used in the ZA.
\citet{Vogelsberger2011} used it in a study of multi-stream structure of galaxy size haloes.
\citet{Shandarin2014a} applied it for identifying subhaloes  in dark matter haloes.
A similar although somewhat simplistic realization of this idea has been revealed in the ORIGAMI method
used for the analysis of the web ( \citealt{Falck2012},  \citealt{Falck2015} ).
 Although these fields cannot be used directly on observational data because the full phase-space information
 is not available, they provide much deeper insight into non-linear clustering of collision-less dark matter
and reveal new features of the web.

@@

We have utilized the data from cosmological N-body simulations by Gadget-2 \citep{Springel2005c} for 100 $h^{-1}$ Mpc  and 200 $h^{-1}$ Mpc box 
sizes with $128^3$, $256^3$ and $512^3$ grids. Each particle is between  $10^9 - 10^{12} M_{\odot}$. 
The initial conditions and cosmological parameters are consistent with the Planck cosmology. We utilize the initial Lagrangian box and do a three-dimensional mapping onto corresponding evolved simulations. In addition, for local multi-stream analyses around haloes, we have utilized halo catalogues for each of these simulation boxes. These haloes are detected using FOF method considering objects with more than 20 particles found at linking length, $ b= 0.2$.  
%-------------------------------------------------------------------------------------------------------------------------------
\section{Multi-stream field calculation}
\label{subsec:method}

@@

\begin{table}
\centering
  \caption{Comparison of the approximate power law dependences of curve fits in Fig. \ref{fig:Vfr_all} and Fig. \ref{fig:Mfr_all}. Power law relations for volume fraction $f_{vol}(n_{str})$ and mass fractions $f_{mass} (n_{str}) $ as a function of number of streams at $n_{str} \geq 5$ are shown (amplitudes are not shown). The boxes of size 100 $h^{-1}$ Mpc on  $128^{3}$ grids and, 200 $h^{-1}$ Mpc on  $256^{3}$ grids have same  $L/N = 0.78 h^{-1}$ Mpc. Similarly, $L/N = 0.39 h^{-1}$ Mpc for boxes of size 100 $h^{-1}$ Mpc on  $256^{3}$ grids and, 200 $h^{-1}$ Mpc on  $512^{3}$ grids.}
\begin{tabular}{|l|r|r|r|r|}
\hline
%\multicolumn{1}{|c|}{} & \multicolumn{4}{l|}{Multi-stream analysis (This work)} & \multicolumn{4}{c|}{ORIGAMI \citep{Falck_Neyrinck:14}}    \\ \hline
$L/N$                & $0.19$ & $0.39$& $0.78$ & $1.56$  \\ \hline
$f_{vol} (n_{str}) $ Vs. $n_{str}$                 & $-2.1$   & $-2.3$    & $-2.5$      & $-2.9$  \\ \hline
$f_{mass} (n_{str}) $   Vs. $n_{str}$            & $-1.1$   & $-1.2$   & $-1.4$       & $-2.0$     \\ \hline
\end{tabular}
 \label{tab:Compare_Slopes}
\end{table}

 \begin{table}
 \centering
  \caption{Comparison of the volume and mass fractions of the void ($n_{str} = 1$) regions of the cosmic web for various simulation boxes at refinement factor of 1. Mean density is the ratio of mass fraction to the volume fraction. It is given in units of the mean density of the universe. The boxes of size 100 $h^{-1}$ Mpc on  $128^{3}$ grids and, 200 $h^{-1}$ Mpc on  $256^{3}$ grids have same  $L/N = 0.78 h^{-1}$ Mpc. Similarly, $L/N = 0.39 h^{-1}$ Mpc for boxes of size 100 $h^{-1}$ Mpc on  $256^{3}$ grids and, 200 $h^{-1}$ Mpc on  $512^{3}$ grids.}
\begin{tabular}{|l|r|r|r|r|}
\hline
%\multicolumn{1}{|c|}{} & \multicolumn{4}{l|}{Multi-stream analysis (This work)} & \multicolumn{4}{c|}{ORIGAMI \citep{Falck_Neyrinck:14}}    \\ \hline
$L/N$                & $0.19$ & $0.39$& $0.78$ & $1.56$  \\ \hline
Volume Fraction (\%)                 & $88$   & $90$    & $93$      & $96$  \\ \hline
Mass Fraction   (\%)               & $24$   & $36$   & $55$       & $77$     \\ \hline
Mean density                    & $0.27$ & $0.40$    & $0.59$        & $0.80$   \\ \hline
\end{tabular}
 \label{tab:Compare_LN}
\end{table}

 \begin{table}
 \centering
  \caption{Comparison of the volume and mass fractions of the void ($n_{str} = 1$) regions of the cosmic web for a simulation box at different refinement factors. Mean density is the ratio of mass fraction to the volume fraction. All the multi-streams for simulation box of length 100 $h^{-1}$ Mpc on raw resolution of $128^{3}$ grids ($ L/N = 0.78 h^{-1}$) }
\begin{tabular}{|l|r|r|r|}
\hline
%\multicolumn{1}{|c|}{} & \multicolumn{4}{l|}{Multi-stream analysis (This work)} & \multicolumn{4}{c|}{ORIGAMI \citep{Falck_Neyrinck:14}}    \\ \hline
Refinement factor               & $1$ & $2$& $8$   \\ \hline
Volume Fraction (\%)                 & $93$   & $93$    & $93$       \\ \hline
Mass Fraction   (\%)               & $55$   & $44$   & $32$           \\ \hline
Mean density                    & $0.59$ & $0.47$    & $0.35$        \\ \hline
\end{tabular}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=10.cm]{Chapter3/Source_v2/fig10a}
\includegraphics[width=10.cm]{Chapter3/Source_v2/fig10b}
\includegraphics[width=6.cm]{Chapter3/Source_v2/fig10c}
\end{minipage}\hfill
\caption{Halo of mass $1.1 \times 10^{12} M_{\odot} $  and FOF radius 0.2 $h^{-1}$ Mpc (dotted-violet line). This halo just has 26 particles, hence the resolution is lesser than previous haloes for surfaces with low radii. Top: There is a bump in fraction of each of 3+ streams a little over 1 $h^{-1}$ Mpc. This is due to the presence of an additional halo nearby, as seen in the projections. The Mollweide projections from 0.4 $h^{-1}$ Mpc to 2 $h^{-1}$ Mpc have high stream flow regions near the lower surfaces of diagnostic spheres. Bottom: Corresponding FOF halo (red, at the centre) has a more massive neighbouring FOF halo (blue) within distance of 2 $h^{-1}$ Mpc. The 17+ stream regions (green) are increased around the neighbouring halo.} 
\label{fig:4322}
\end{figure}

@@

 \begin{table*}
 \centering
  \caption{Comparison of the volume and mass fraction of the elements of the cosmic web between our analysis and \citet{Falck2015}. $ L/N = 0.78 h^{-1}$ Mpc for simulations used in both techniques. We use a refinement factor of 8 for the multi-stream grid. The mean density is given in units of the mean density of the universe. }
\begin{tabular}{|l|r|r|r|r|r|r|r|r|}
\hline
\multicolumn{1}{|c|}{} & \multicolumn{4}{l|}{Multi-stream analysis (This work)} & \multicolumn{4}{c|}{ORIGAMI \citep{Falck2015}}    \\ \hline
                       & Voids & Walls & Filaments & Haloes  & Voids & Walls & Filaments & Haloes \\ \hline
Volume Fraction (\%)                 & $93$   & $7$    & $< 1$      & $< 0.1$ & $84$   & $12$   & $3$        & $< 1$  \\ \hline
Mass Fraction   (\%)               & $32$   & $35$   & $17$       & $14$    & $26$   & $19$   & $19$       & $35$   \\ \hline
Mean density                    & $0.34$ & $5$    & $>17$      & $>140$  & $0.31$ & $2.2$  & $6$        & $>35$  \\ \hline
\end{tabular}
 \label{tab:Compare}
\end{table*}

@@

% \section*{Acknowledgements}
% The authors are grateful to Gustavo Yepes and Noam Libeskind for providing the simulations used in this study and the Sir John Templeton Foundation for the support.The authors would also like to thank the anonymous referee for careful review and useful suggestions.
